Social Visual Analytics: DevOps for Data Scientists

### 2015-02-18 introduction.tex
yes

Consider the role of a data science team within an organization today.
Analysts work in loose or tight collaboration with colleagues.
Project teams vary in size, from just a few people to dozens or
more, even within one project's duration. Tasks are broad, including problem
identification, data wrangling, visualization, modeling and analysis,
interpreting, explaining and advocating the results, and often helping
clients to realize the benefits of the work in ongoing processes.
There are many details to these tasks.
For example, data wrangling can involve finding data, figuring out its
syntax and semantics, assessing data quality, performing normalization and
data quality remediation, and making the data available for other tasks
to follow.

Visual analytics in practice depends greatly on communication and collaboration.
At almost every step, detailed knowledge about data, code and tasks
is shared by collaborators. 
Further, data scientists are increasingly asked to work more closely
with business people who are domain specialists and less technically oriented.
Thus, data scientists and developers are being asked to become
very broad, and integrate work across the spectrum. (Hiring ``full stack
developers'' to perform integration seems to be a trend, but does not achieve
the potential benefits of specialization and division of labor in larger
teams, and is not a practice to be encouraged.)
This pushes the boundaries for communication and collaboration even further.
Unfortunately, processes and technology that support this work are fragmented.
Knowledge is shared through informal conversations, email online messages,
meetings, phone calls, within code itself, in wikis, in project documents and
by other means. Results of experiments are often shared first by copying static
output, later, by holding meetings that lead to moving or even rewriting
entire programs.

>>>>

Over fifty years ago, John Tukey published ``The Future of Data Analysis''
in which he defined a remarkably prescient agenda for the future of
data analysis and statistical computing. His vision has been realized
through environments like S and R, that started  a revolution by
integrating statistical graphics, computation, and robust techniques
for practical data modeling. Every day, many thousands of data scientists
and others specialists benefit from this brilliant vision.

The field has now moved beyond the boundaries of what Tukey originally
foresaw. Processing and networking capabilities today far exceed
what was barely imagined fifty years ago. Inexpensive, scalable
virtualized systems allow analysts to routinely solve problems
involving millions or even billions of data values. Networking
also enables both collaboration and the automation of complex
tasks and processes in ways that were not considered even ten years ago.
We turn our attention to this opportunity to use computation
to support, not only individuals, but entire teams and their
work within larger organizations.

As a starting point, consider the role of a data science team within
a business or technical organization today.
Analysts work in loose or tight collaboration with colleagues and
domain experts.
Project teams vary in size, from just a few people to dozens or
more, even within one project's duration. Assignments are often
broad, and include tasks such as problem identification,
data wrangling, modeling, analysis, visualization, summarization
and interpretation of results, and explaining them and recommending
decisions, helping clients to realize the benefits of the analysis.
Eventually, knowledge or working prototypes created by data science teams
may be transferred to other organizations to employ them in production.
There are many details to these tasks. For example, data wrangling can involve
finding data, understanding its syntax and semantics, assessing data quality,
performing normalization and data quality remediation, and making the data
available to other tasks that will follow.

Visual analytics depends greatly on communication and collaboration.
At almost every step, detailed knowledge about data, code and tasks
is shared by collaborators. 
Further, data scientists are increasingly asked to work more closely
with business people who are domain specialists and less technically oriented.
Thus, data scientists and developers are being asked to become
very broad, and integrate work across the spectrum.

For example, interview comments found in {\it Data Scientists at Work}
by Sebastian Gutierrez illuminate the current practice of visual analytics.
There is a common theme, that data science or visual analytics is 
being done in teams that more share their work, both data and code,
more freely than before. We particularly want to point out the remarks
by Jonathan Lenaghan, who anticipates an environment for continuous
integration and publication of data analysis.

\begin{addmargin}[1em]{1em}  % left and right margin change
\small
\input{quotes}
\end{addmargin}

% (Hiring ``full stack developers'' to perform integration seems to be a trend, but does not achieve the potential benefits of specialization and division of labor in larger teams, and is not a practice to be encouraged.)

Unfortunately, processes and technology that support visual analytics today
work are fragmented.
Knowledge is shared through informal conversations, email online messages,
meetings, phone calls, within code itself, in wikis, in project documents and
by other means. Results of experiments are often shared first by copying static
output, later, by holding meetings that lead to moving or even rewriting
entire programs.

### 2015-02-18 quotes.tex
yes

% part1.txt
Shellman: Finally, prototyping our products so that internal
customers can use them early on has been crucial for
our success. It doesn't even have to be something
fancy. For instance, our recommendations preview
tool doesn't have a particularly interesting
visualization, but its enough to show the result of our
algorithms. Now we can shoot off a URL to internal
customers and it allows them to sit at their desk and
understand the behavior of our product and
experiment with it, and provide feedback way before
we're talking about getting it into production. This
has been super helpful and has been a really great
way to get people excited about what we're working
on. Building and maintaining those relationships are
just like anything else-- you always have to be
working on them. Nurturing those relationships are
part of the job, and if you leave them unattended,
they might not be there later.

@@

% part3.txt
Foreman: I lead the data science team at
MailChimp, and I like to get my hands dirty too.
Some of the big pieces of my day involve working
with my team to take stock of current projects and
figure out where to go next, doing my own work and
prototyping things-- I do projects just like my peers
do and then also my talking with other teams,
talking with management, and planning for the
future.
On our team, we’ve got different folks facing different
% part4.txt
kinds of projects. We’ve got one person who really
owns compliance and looks at our compliance
processes. We’ve got another data scientist who
focuses on more of the user experience side of the
house and understanding our customers. I help them
and others as is needed while I do the three things
mentioned earlier-- build data products, be a
translator, and have conversations with the data
science community.

% part5.txt
Lenaghan: First and foremost, it is very important
to be self-critical: always question your assumptions
and be paranoid about your outputs. That is the easy
part, In terms of skills that people should have if they
really want to succeed in the data science ﬁeld, it is
essential to have good software engineering skills. So
even though we may hire people who come in with
very little programming experience, we work very
hard to instill in them very quickly the importance of
engineering, engineering practices, and a lot of good
agile programming practices, This is helpful to them
and us, as these can all be applied almost one-to-one
to data science right now.
If you look at dev ops right now, they have things
such as continuous integration, continuous build,
automated testing, and test harnesses-- a1l of which
map very well from the dev ops world to the data ops
(a phrase I stole from Red Monk) world very easily. I
think this is a very powerful notion. It is important to
have testing frameworks for all of your data, so that if
you make a code change, you can go back and test all
of your data. % Having an engineering mindset is

% part6.txt
In addition to Bitly’s extraordinary data sets, I was
given all the resources that I needed. These were
resources that I’d only heard about but had never
seen. I went from asking what a Hadoop cluster was,
as my school didn't have one, to being able to work
with one. Now I could play with one to figure out how
to make my work better and how to change my
algorithms so they were cleaner and ran faster on
Hadoop. The whole transition was eye opening. Not
only that, I could also look at other people’s code and
play with their code and data sets as well. It was
fantastic and I was convinced that working with data
was my thing.

>>>>

% part1.txt
Erin Shellman: Finally, prototyping our products so that internal
customers can use them early on has been crucial for
our success. It doesn't even have to be something
fancy. For instance, our recommendations preview
tool doesn't have a particularly interesting
visualization, but its enough to show the result of our
algorithms. Now we can shoot off a URL to internal
customers and it allows them to sit at their desk and
understand the behavior of our product and
experiment with it, and provide feedback way before
we're talking about getting it into production. This
has been super helpful and has been a really great
way to get people excited about what we're working
on. Building and maintaining those relationships are
just like anything else-- you always have to be
working on them. Nurturing those relationships are
part of the job, and if you leave them unattended,
they might not be there later.

@@

% part3.txt
John Foreman: I lead the data science team at
MailChimp, and I like to get my hands dirty too.
Some of the big pieces of my day involve working
with my team to take stock of current projects and
figure out where to go next, doing my own work and
prototyping things-- I do projects just like my peers
do and then also my talking with other teams,
talking with management, and planning for the
future.
On our team, we’ve got different folks facing different
% part4.txt
kinds of projects. We’ve got one person who really
owns compliance and looks at our compliance
processes. We’ve got another data scientist who
focuses on more of the user experience side of the
house and understanding our customers. I help them
and others as is needed while I do the three things
mentioned earlier-- build data products, be a
translator, and have conversations with the data
science community.

% part5.txt
Jonathan Lenaghan: First and foremost, it is very important
to be self-critical: always question your assumptions
and be paranoid about your outputs. That is the easy
part, In terms of skills that people should have if they
really want to succeed in the data science field, it is
essential to have good software engineering skills. So
even though we may hire people who come in with
very little programming experience, we work very
hard to instill in them very quickly the importance of
engineering, engineering practices, and a lot of good
agile programming practices, This is helpful to them
and us, as these can all be applied almost one-to-one
to data science right now.
If you look at dev ops right now, they have things
such as continuous integration, continuous build,
automated testing, and test harnesses-- a1l of which
map very well from the dev ops world to the data ops
(a phrase I stole from Red Monk) world very easily. I
think this is a very powerful notion. It is important to
have testing frameworks for all of your data, so that if
you make a code change, you can go back and test all
of your data. % Having an engineering mindset is

% part6.txt
Anna Smith: In addition to Bitly’s extraordinary data sets,
I was given all the resources that I needed. These were
resources that I’d only heard about but had never
seen. I went from asking what a Hadoop cluster was,
as my school didn't have one, to being able to work
with one. Now I could play with one to figure out how
to make my work better and how to change my
algorithms so they were cleaner and ran faster on
Hadoop. The whole transition was eye opening. Not
only that, I could also look at other people’s code and
play with their code and data sets as well. It was
fantastic and I was convinced that working with data
was my thing.
